Classless Java

### 2015-09-25 paper/sections/Overview.tex
revised overview, with subsection structure

\section{Overview}\label{sec:ep}
Yanlin

* Explain what the Mixin annotations do using examples.

* Motivate the use of multiple inheritance in Java.

* Maybe use Marco's example (Point example).

We provide a Java annotation \mixin to provide default implementations for
various methods and a mechanism to instantiate objects. \mixin annotation helps
programmers to write less cumbersome code and instantiate interfaces in
Java. More over, with \mixin annotation, programmers can mimic multiple
inheritance in a restricted form easily.

For example, interface \texttt{Point} annotated with \mixin:

@@

\texttt{Point} has two (conceptually) member fields \texttt{X} and
\texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
methods. % Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
% \emph{setter} methods. Method \texttt{Point withX(int X)} updates field
% \texttt{X} and returns \textbf{this}.

A typical and trivial implementation that programmers usually do is:

\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
may not be reused in a single inheritance language. However, with our approach,
the \mixin annotation will generate a static method \texttt{of} inside
\texttt{Point}. \texttt{of} makes use of Java anonymous classes and achieves the
same implementation as \texttt{PointImpl}.

@@

Apart from generating code for \emph{getter} methods, the annotation processor
can also generate \emph{setter} methods, \emph{with} methods, \texttt{clone}
methods, etc. We summarize all the forms of method \mixin  supports below. Inside
the anonymous class in the annotated interface, the following code are
generated:

@@

\item For methods inside the interface with the form \texttt{void x(Tx x)}:
  \begin{itemize}
    \item check if exist method \texttt{Tx x()}. If not, generate error.
    \item generate implemented setter method:
       \begin{lstlisting}
           pubic void x(Tx x) { this.x = x; }
       \end{lstlisting}
    \end{itemize}

@@

\item For methods with the form of \texttt{T x(Tx \_)}:
  \begin{itemize}
    \item check if exist method \texttt{T x()}, if not, generate error.
    \item inside the inner class, generate
      \begin{lstlisting}
          pubic T x(Tx x) { this.x = x; return this;}
      \end{lstlisting}
  \end{itemize}
\end{itemize}

Another advantage: multiple inheritance.

>>>>

\section{Overview}\label{sec:ep}

\subsection{A Running Example: \texttt{Point}}
Suppose we want to create a point component that models the a point in space,
that has x-coordinate and y-coordinate. For example, if we create the
\texttt{Point} interface in Java, it would look like this:

\begin{lstlisting}
interface Point {
    int x();
    int y(); 
}
\end{lstlisting}

\texttt{Point} has two (conceptually) member fields \texttt{x} and \texttt{y},
representing the two coordinates of a point. Methods \texttt{int x()} and
\texttt{int y()} serve as \emph{getter} methods. 
% Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
% \emph{setter} methods. Method \texttt{Point withX(int X)} updates field
% \texttt{X} and returns \textbf{this}.

\subsection{Naive Implementation}
In Java, to implement an interface like \texttt{Point}, a typical and trivial
approach that programmers usually do is creating a class extending the interface
and providing implementation for all methods inside. For example, this is the
implementation for interface \texttt{Point}:

\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
may not be reused in a single inheritance language.

\subsection{Our Approach}
Instead of writing a whole another class to provide the implementation for
\texttt{Point}, we annotate on interface \texttt{Point} directly with \mixin:

@@

The \mixin annotation will generate a static method \texttt{of} inside
\texttt{Point}. The method \texttt{of} mimic the functionality of constructors,
it takes arguments same as constructors and return objects similar to
constructors. It makes use of Java anonymous classes and achieves the same
implementation as \texttt{PointImpl}.

With our approach, we provide a Java annotation \mixin to provide default
implementations for various methods and a mechanism to instantiate
objects. \mixin annotation helps programmers to write less cumbersome code and
instantiate interfaces in Java.

@@

\subsection{What  \mixin Generates}
Apart from generating code for \emph{getter} methods, the annotation processor
can also generate \emph{setter} methods, \emph{with} methods, \texttt{clone}
methods, etc. We summarize all the forms of method \mixin  supports below. Inside
the anonymous class in the annotated interface, the following code are
generated:

@@

\item For methods inside the interface with the form \texttt{void x(Tx x)}:
  \begin{itemize}
    \item check if exist method \texttt{Tx x()}. If not, generate error.
    \item generate implemented setter method:
       \begin{lstlisting}
           public void x(Tx x) { this.x = x; }
       \end{lstlisting}
    \end{itemize}

@@

\item For methods with the form of \texttt{T x(Tx \_)}:
  \begin{itemize}
    \item check if exist method \texttt{T x()}, if not, generate error.
    \item inside the inner class, generate
      \begin{lstlisting}
          public T x(Tx x) { this.x = x; return this;}
      \end{lstlisting}
  \end{itemize}
\end{itemize}

\subsection{More On \mixin}
Besides the benefit of freeing programmers from writing boilerplate code, our
\mixin annotation can also allow programs to mimic multiple inheritance in a
restricted form easily.
